---
title: "Initial Korean covid 19"
author: "Karen Gomez"
date: "Summer 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```



## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + sex,data=kRdata)
```

In this data there are more women than men alive. They outnumber the males significantly as shown by the blue bar in the frequency chart above. The chart also shows that in this group people there are more people alive than dead as shown by the two bars labeled alive on the left and the two bars labeled dead on the right.

```{r}
barchartGC(~result + sex,data=kRdata, type="percent")

``` 

In this percent chart we see that the two bars labeled "alive" are not the same as the two bars labeled "dead". The two bars to the left show that there are more women alive than man whereas the bars to the right show that there are more men dead than women. The percent of men dead is close to 63 percent and the percent of women dead is around 37 percent. Representing that there is a significant difference between the "alive" bars and "dead" bars. Meaning that an unexpected situation occurred because if everything were normal both alive and dead bars would be similar.
  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The first part of the numerical data shows how many females and males were alive which was 56% and 44% respectively. It also shows how many females and males were dead which was 36% and 64% respectively. This portion represents the rows of the chart shown previously. In the second part of this data it shows the result of what percent of the female population were dead which was 1.4% and the result of what percent of the male population was dead which was 3%. This demonstrates that the population of dead males was more than twice bigger than the population of the dead females. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

```{r}
fisher.test(table1)
```

In this fisher test we will separately compute the odds ratio. The odds of a female dying is 25/1831 and the odds of a man dying is 44/1411. So, the odds ratio is (44/1411)/(25/1831) which equals to 2.3. So, we reject the null hypothesis because we have a significant results. There is a probability of 1 in 1000 times to ever have something like this happen.

## Conclusions

For the first part of this problem we first had to upload our categorical and categorical helper file because the data presented had two factor variables. Then we ran the frequency plot and the percent plot showing us the difference between the female and male population that are alive versus the female and male population that are dead. Then we ran the numerical results which showed us the exact percentage of dead females and the exact percentage of dead males. Which was 1.4% for females and 3% for males. Demonstrating that the population of dead males was more than twice higher than the percent of dead females. Then we ran the inferential test which showed us the chi square test and the fisher test, demonstrating that there is a probability of 1 in 1000 that this situation would ever occur. 
